---
title: 'Demo: simulation function and visualizing results'
author: "J. Kirby and M. Sonderegger"
date: '2018-03-16'
output: 
    html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This file explains how to carry out the kind of analysis shown in the main paper, in more user-friendly fashion than the code.  Code is illustrated from:

* `runSimsJPhon.R` : run the exact simulations in the paper

* `powerEs.R`:functions to carry out simulations used to approximate power and Type M/S error

* `summarizeVisualizeRuns.R`: functions to summarize and visualize results 

# 1. Run simulations

Load and process Roettger et al. data:
```{r, message=FALSE, warning=FALSE}
library(lme4)
library(arm)
library(plyr)
library(dplyr)
library(ggplot2)
library(simr)


E1 = read.csv("roettgerEtAlData.csv",comment.char="")
E1 <- E1[-481,]

## turn items, subjects into factors
E1$item_pair = as.factor(E1$item_pair)
E1$subject = as.factor(E1$subject)

## center voicing variable
E1$voiceless <- rescale(E1$voicing)
```

Fit a mixed model to the R2014 data:
```{r}
mod  <- lmer(vowel_dur ~ voiceless  + accent_type + prosodic_boundary + place + vowel + norming_voiceless_count + (1+voiceless|subject) + (1+voiceless|item_pair),
              data=E1)
```

These two things are done by executing 
```{r}
source('powerEs.R')
```

which also contains definitions of the functions `sampleNewDataset`, `runSim`, and `runSweep`.

Each call to `runSim` runs a simulation for fixed parameter values. For example, to run a simulation with `nSim=100` for the actual parameter values for the Roettger et al. (2014) data:
```{r, cache=TRUE, warning=FALSE}
useParallel <- FALSE
runs <- runSim(E1.mdl, nS=16, nI=24, nRep=1, beta=-9, nSims=100)
```

Each row of the resulting dataframe corresponds to a single model, fitted to simulated data:
```{r}
head(runs)
```

The `runSweep` function runs simulations for a range of parameter values (one simulation = one call to `runSim`). For example, to carry out simulations across the parameter values shown in Table 1 of the paper, with `nSims=1500`, you use the code in `runSimsJphon.R`:
```{r, eval=FALSE, echo=TRUE}
useParallel <- TRUE
nCores <- 25

if(useParallel){
    library(doMC)
    registerDoMC(nCores)
    getDoParWorkers()
}
runSweep(mod=origMod, betaVals=seq(-2, -10),
         nSVals <- seq(6,26,4),
         nIVals <- seq(10,30,5),
         nRepVals <- seq(1,6),
         nSims=1500,
         saveFName = 'runs_16Mar18_nRuns1500.rds')
```

Note that:

* The simulation sweeps automatically execute in parallel, if you specify a number of  cores and use the `useParallel` option

* The code directly above will take a very long time (days), so it is not recommended to execute it on a laptop. Instead, we execute it on a remote server, and save the resulting runs as an RDS file.


# 2.  Summarize and visualize results

To load in the the results from an RDS file:
```{r}
runs <- readRDS("runs_16Mar18_nRuns1500.rds")
```

The file `summarizeVisualizeRuns.R` contains several functions to calculate power and Type M/S error based on these runs:
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)

binomEst <- function(n1, n){
    if(n>0){
        temp <- binom.test(n1, n)
        return(c(temp$conf.int[1], temp$conf.int[2], as.numeric(temp$estimate)))
    } else{
        return(c(NA, NA, NA))
    }
}

normalEst <- function(x){
    m <- mean(x, na.rm=TRUE)
    se <- sd(x, na.rm = TRUE)/sqrt(length(which(!is.na(x))))
    return(c(m, m-1.96*se, m+1.96*se))
}

## function to compute summary df for power -- estimate, and lower and upper bounds
## for 95% CI --
## as a function of all input vars, just using LR model selection strategy
##
powerSummary <- function(x){
    
    powerVals <- function(nSig, n){
        temp <- binom.test(nSig, n)
        return(c(temp$conf.int[1], temp$conf.int[2], as.numeric(temp$estimate)))
    }
    
    summaryDf <- x %>% group_by(nS,nI,nRep, trueBeta) %>% 
        dplyr::summarise(nSig=sum(lrTestModSelect=='yes'), n=n()) %>% 
        group_by(nS,nI,nRep, trueBeta, n, nSig) %>% 
        do(setNames(data.frame(t(powerVals(.$nSig, .$n))), c('power_lower', 'power_upper', 'power')))

    return(summaryDf)                                         
    
}



## make plottable dataframe for just for just typeM and typeS error,
## broken down by whether conditional on significance, non-significance, or using both.
typeMSSummary <-  function(x){
    
    
    
    ## summary for typeM
    tempDf1 <- x %>% mutate(betaRat = abs(beta)/abs(trueBeta), sameSign=(sign(beta)==sign(trueBeta))) %>%
        group_by(nS,nI,nRep, trueBeta) %>%
        do(
            cbind(
                setNames(data.frame(t(normalEst(.$betaRat))),
                         c('typeM_both_est', 'typeM_both_lower', 'typeM_both_upper')),
                setNames(data.frame(t(normalEst(ifelse(.$lrTestModSelect=='yes',.$betaRat,  NA)))), 
                          c('typeM_sig_est', 'typeM_sig_lower', 'typeM_sig_upper')),
                setNames(data.frame(t(normalEst(ifelse(.$lrTestModSelect=='no',.$betaRat,  NA)))), 
                         c('typeM_ns_est', 'typeM_ns_lower', 'typeM_ns_upper'))
            )
        )
                         
              
    tempDf2 <- x %>% mutate(sameSign=(sign(beta)==sign(trueBeta))) %>%
        group_by(nS,nI,nRep, trueBeta) %>%
        dplyr::summarise(nSig=sum(lrTestModSelect=='yes'), n=n(),
                  nSigRightSign=sum(sameSign & lrTestModSelect=='yes'),
                  nNsRightSign=sum(sameSign & lrTestModSelect=='no'),
                  power = nSig/n()) %>%
        group_by(nS,nI,nRep, trueBeta, power) %>%
        do(
            cbind(
                setNames(data.frame(t(binomEst(.$n-.$nSigRightSign -.$nNsRightSign, .$n))), c('typeS_both_lower', 'typeS_both_upper', 'typeS_both_est')),
                setNames(data.frame(t(binomEst(.$nSig-.$nSigRightSign, .$nSig))), c('typeS_sig_lower', 'typeS_sig_upper', 'typeS_sig_est')),
                setNames(data.frame(t(binomEst(.$n-.$nSig-.$nNsRightSign, .$n-.$nSig))), c('typeS_ns_lower', 'typeS_ns_upper', 'typeS_ns_est'))
            )
        )
     tempDf <- left_join(tempDf1, tempDf2)   
    
    tempDf <- tempDf %>% gather("key", "val", starts_with("type")) %>%
        separate(key, into=c('var','sig','measure')) %>% 
        unite(measure, var, measure) %>% 
        spread(measure, val)
    return(tempDf)
    
}
```

For example, let's consider just the "low sample size" case from the paper: 6 subjects, 10 items, 1 repetition, for a range of effect sizes:

```{r}
lowSubset <- filter(runs, nS==6 & nI==10 & nRep==1)
```

To calculate power and Type M/S error using these simulation runs, for each unique value of the effect size:
```{r}
powerSumm <- powerSummary(lowSubset)
typeMsSumm <- typeMSSummary(lowSubset)
```

For example, the power dataframe:
```{r}
powerSumm
```

shows that the estimated power for effect size -8 is 0.273, with 95% CI (0.251, 0.297).

We can now plot power as a function of effect size (like Fig. 4 in the paper), showing a dotted line at the customary 80% cutoff value:
```{r}
ggplot(aes(x=trueBeta, y=power), data=powerSumm) + 
    geom_line() +
    geom_ribbon(aes(ymin=power_lower, ymax=power_upper), alpha=0.25) + geom_hline(yintercept=0.8, lty=2) + ylim(0,1) +
    xlab("True effect size (ms)") + ylab("Power") + geom_hline(yintercept=0.8, lty=2) + theme_set(theme_bw()) + ggtitle("Power plot")
```


We can also plot Type M and Type S error as a function of effect size (like Figs. 5-6 in the paper), using separate panels to show values conditioned on significance, or calculated across all runs:
```{r}
 ggplot(aes(x=trueBeta, y=typeM_est), data = typeMsSumm) + 
    geom_line() + 
    geom_ribbon(aes(ymin=typeM_lower, ymax=typeM_upper), alpha=0.25) +
    xlab("True effect size") + ylab("Type M error")  +
    geom_hline(yintercept=1, lty=2) +  theme_set(theme_bw())  + 
    facet_wrap(~sig) + theme_set(theme_bw()) + ggtitle("Type M error plot")

```


```{r}
ggplot(aes(x=trueBeta, y=typeS_est), data = typeMsSumm) + 
    geom_line() + 
    geom_ribbon(aes(ymin=typeS_lower, ymax=typeS_upper), alpha=0.25) + theme_set(theme_bw())  + 
    xlab("True effect size") + ylab("Type S error") + xlab("True effect size (ms)")  + geom_hline(yintercept=0, lty=2) + 
    facet_wrap(~sig) +  theme_set(theme_bw()) + ylim(0,1) + ggtitle("Type S error plot")

```
